The problem and promise of scale in multilayer networks

### 2020-02-03 paper/manuscript.Rmd
network metrics

# Abstract
Scale remains a seminal concept in ecology. Spatial scale, for instance, has
become a central consideration in the way we understand landscape ecology and
animal space use. Meanwhile, social processes can also scale from fine-scale
interactions to co-occurrence to overlapping home ranges. Multilayer networks
promise the integration of monolayer animal social networks with the complexity
and importance of animal space use and movement in heterogeneous landscapes.
Despite the complex interplay between social networks and how animals use space,
there remains an important biological and methodological gap in our
understanding of an animal’s perception of scale, e.g., grain and extent. But
scale can affect multilayer network dynamics and how we build and interpret
them. Here, we discuss the role of scale in the context of multilayer networks
and provide a case study of caribou (*Rangifer tarandus*) to illustrate the role
of social,spatial, and temporal scale on multilayer processes. We also present
perspectives on future development and application of scale in multilayer
networks with respect to habitat selection and space use, individual fitness and
phenotypes, and movement ecology and collective movement.Effective integration
of social and spatial processes, including biologically meaningful scales,
within the context of animal social networks is an emerging area of research;
our contribution uniquely incorporates perspectives on how the social
environment and spatial processes are linked across scales in a multilayer
framework. Based on social network and behavioural ecology theory as well as our
case study, we demonstrate potential for interconnectedness among individual
animals, their social environment, and the landscapes they occupy.

# Introduction
Our inference regarding the ecological processes underlying ecological patterns
are directly linked to the scale at which they are observed [@Levin_1992;
@Allen_2015]. Animals are influenced across scales, through processes such as
species-specific as trophic interactions and resource availability
[@Legendre_1993; @Chave_2013] and global climate and productivity [@Field_2009].
Multiscale ecology has been integrated into species distribution modelling
[@Elith_2009], habitat selection [@Mayor_2007], and food webs [@Sugihara_1989].
Meanwhile, in the context of animal behaviour, certain behaviours are scale
dependent, for example, acoustic communication in birds occurs over variable
spatial scales [@Luther_2009]. We posit that animal behaviour and, more
specifically, the social environment that animals experience must also be
scale-dependent.

@@

# Scale in multilayer networks: a case study

## Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics [@Lesmerises_2018] that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage (Bergerud 1974)<!-- TODO-->. As a result, caribou tend to occupy
and re-use craters once they are established and there is considerably less
access to forage than when the landscape is snow-free. Importantly, craters
exist on the landscape at multiple scales: the crater scale (a single crater),
the feeding area scale (multiple craters in close proximity), and the winter
range scale (all craters within an individual’s range) [@Mayor_2009]. The
distribution of forage for caribou is therefore seasonally heterogeneous, with
greater access in snow-free seasons compared to winter as well as spatially
heterogeneous in winter because the distribution of craters on the landscape
varies meaning that access to vegetation in winter is highly variable for
caribou. We therefore partitioned data into two discrete seasons:  summer (3
August - 20 September) and winter (1 January – 17 February) based on caribou
socioecology [@Peignier_2019].

## Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas
et al. (2010)<!-- TODO-->. We did not collar all female caribou in the herd;
however, we assumed that our sample of collared animals was random.

## Landscape data and habitat classification
Landcover data were provided by the Newfoundland and Labrador Wildlife Division.
Available landcover classification included nine habitat types at <!-- TODO:
double check --> 10m resolution. We reclassified the landcover types into three
categories: foraging habitat, open habitat, and forest habitat. Foraging habitat
<!-- or just lichen like figs?--> consisted of lichen habitat, open habitat
consisted of wetland, rocky barrens, and anthropogenic habitat types, while
forest habitat consisted of coniferous forest, conifer scrub, broadleaf forest,
and mixed-wood forest habitat types. Water habitat was excluded from all
subsequent analyses.
<!-- using the raster package for R? -->

## Caribou multilayer social networks

### Network types and construction (defining layers for the multilayer network)
<!-- should use something like "using syntax from Kivela 2014..." or 
Following [@Kivela_2014]:
other info to maybe include: the multilayer network was not node-aligned since
not all individuals were on all layers
the temporal aspect is ordinal coupling and landcover aspect is categorical coupling

@@

### Network metrics
We used a series of metrics to characterize the multilayer networks focusing on
the relevance of individual layers, similarity of layers and the role of
individuals. We calculated multidegree, the sum of degree centrality across
layers, and neighbourhood, the sum of unique neighbours to each individual
[@Berlingerio_2012; @Kivela_2014]. Comparing the multidegree and number of
unique neighbours, we measured the connective redundancy, the proportion of
neighbours repeated in multiple layers, and the layer relevance, the proportion
of neighbours present compared to the neighbourhood for each individual
[@Berlingerio_2012]. To measure the role of individuals, we calculated the graph
strength for each individual for each layer specific season and landcover
context <!-- citation?-->. Finally, we calculated the similarity of layers
between seasons, across landcover classes (e.g.: winter open compared to summer
open) using the Pearson correlation coefficient [@Br_dka_2018].

## Varying scale in multilayer networks
### Social scale

@@

### Spatial scale

@@

### Temporal scale

@@

# Scale in multilayer networks in practice: case study results and discussion

## Each metric's interpretation
* Multidegree: sum of degree across layers. High multidegree = connected to many
individuals across layers. Low multidegree = not connected to many individuals
across layers.

@@

## Figure captions and brief interpretations

@@

(ref:lcres) Variation in multilayer network metrics as a function of changes in landcover resolution, a metric which we interpret as spatial scale. For each landcover resolution (10m, 100m, 250m, 500m and 1000m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter), and four network metrics were calculated and we examined how network metrics changed as a function of changes in spatial scale. A) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an increase in similarity for forest layers at 250m and decrease in similarity for open layers at 250m. Note that lichen layer similarity is not shown because individuals did not interact in lichen habitats greater greater than 10m landcover resolution. B) Mean connective redundancy across individuals showed a decreasing trend with increasing landcover resolution. C) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter lichen layer relevance at higher landcover resolutions. D) Mean graph strength of individuals was weaker during the summer compared to the winter months.

@@

##
Results and interpretation of temporal scale

@@

# Perspectives  to advance multilayer networks 
## Habitat selection and space use

@@

## Individual fitness and phenotypes

@@

## Movement ecology and collective movement

@@

# Conclusions

@@

# Acknowledgements

We thank all members of the Wildlife Evolutionary Ecology Lab, including J. Hendrix, K.
Kingdon, S. Boyle,  C. Prokopenko, I. Richmond, J. Hogg, and L. Newediuk for
their comments on previous versions of this manuscript as well as D. C. Schneider
for inspiration and helpful discussions about scale. Funding for this study was
provided by a Vanier Canada Graduate Scholarship to QMRW and a NSERC Discovery
Grant to EVW.

# References

>>>>

## Abstract
Scale remains a seminal concept in ecology. Spatial scale, for instance, has
become a central consideration in the way we understand landscape ecology and
animal space use. Meanwhile, social processes can also scale from fine-scale
interactions to co-occurrence to overlapping home ranges. Multilayer networks
promise the integration of monolayer animal social networks with the complexity
and importance of animal space use and movement in heterogeneous landscapes.
Despite the complex interplay between social networks and how animals use space,
there remains an important biological and methodological gap in our
understanding of an animal’s perception of scale, e.g., grain and extent. But
scale can affect multilayer network dynamics and how we build and interpret
them. Here, we discuss the role of scale in the context of multilayer networks
and provide a case study of caribou (*Rangifer tarandus*) to illustrate the role
of social,spatial, and temporal scale on multilayer processes. We also present
perspectives on future development and application of scale in multilayer
networks with respect to habitat selection and space use, individual fitness and
phenotypes, and movement ecology and collective movement.Effective integration
of social and spatial processes, including biologically meaningful scales,
within the context of animal social networks is an emerging area of research;
our contribution uniquely incorporates perspectives on how the social
environment and spatial processes are linked across scales in a multilayer
framework. Based on social network and behavioural ecology theory as well as our
case study, we demonstrate potential for interconnectedness among individual
animals, their social environment, and the landscapes they occupy.

## Introduction
Our inference regarding the ecological processes underlying ecological patterns
are directly linked to the scale at which they are observed [@Levin_1992;
@Allen_2015]. Animals are influenced across scales, through processes such as
species-specific as trophic interactions and resource availability
[@Legendre_1993; @Chave_2013] and global climate and productivity [@Field_2009].
Multiscale ecology has been integrated into species distribution modelling
[@Elith_2009], habitat selection [@Mayor_2007], and food webs [@Sugihara_1989].
Meanwhile, in the context of animal behaviour, certain behaviours are scale
dependent, for example, acoustic communication in birds occurs over variable
spatial scales [@Luther_2009]. We posit that animal behaviour and, more
specifically, the social environment that animals experience must also be
scale-dependent.

@@

## Scale in multilayer networks: a case study

### Caribou socioecology
Here, we use caribou as a case study to develop scale-dependent multilayer
networks based on social (visual and auditory perception), spatial (habitat
type), and temporal (seasonal). Caribou are gregarious ungulates with
fission-fusion dynamics [@Lesmerises_2018] that display temporal
[@Peignier_2019] and spatial (Webber et al. unpublished data) variation in
social networks. In winter, caribou dig holes in the snow, termed craters, to
access forage (Bergerud 1974)<!-- TODO-->. As a result, caribou tend to occupy
and re-use craters once they are established and there is considerably less
access to forage than when the landscape is snow-free. Importantly, craters
exist on the landscape at multiple scales: the crater scale (a single crater),
the feeding area scale (multiple craters in close proximity), and the winter
range scale (all craters within an individual’s range) [@Mayor_2009]. The
distribution of forage for caribou is therefore seasonally heterogeneous, with
greater access in snow-free seasons compared to winter as well as spatially
heterogeneous in winter because the distribution of craters on the landscape
varies meaning that access to vegetation in winter is highly variable for
caribou. We therefore partitioned data into two discrete seasons:  summer (3
August - 20 September) and winter (1 January – 17 February) based on caribou
socioecology [@Peignier_2019].

### Caribou location data
We used GPS location data collected from Fogo Island caribou in 2017 and 2018.
Adult female caribou were immobilized and fitted with global positioning system
(GPS) collars (Lotek Wireless Inc., Newmarket, ON, Canada, GPS4400M collars,
1,250 g) as described by Schaefer and Mahoney (2013). Collars were programmed to
collect location fixes every 2 hours. Prior to analyses, we subset GPS fixes
into a discrete removed all erroneous and outlier GPS fixes following Bjørneraas
et al. (2010)<!-- TODO-->. We did not collar all female caribou in the herd;
however, we assumed that our sample of collared animals was random.

### Landscape data and habitat classification
Landcover data were provided by the Newfoundland and Labrador Wildlife Division.
Available landcover classification included nine habitat types at <!-- TODO:
double check --> 10m resolution. We reclassified the landcover types into three
categories: foraging habitat, open habitat, and forest habitat. Foraging habitat
<!-- or just lichen like figs?--> consisted of lichen habitat, open habitat
consisted of wetland, rocky barrens, and anthropogenic habitat types, while
forest habitat consisted of coniferous forest, conifer scrub, broadleaf forest,
and mixed-wood forest habitat types. Water habitat was excluded from all
subsequent analyses.
<!-- using the raster package for R? -->

### Caribou multilayer social networks

#### Network types and construction (defining layers for the multilayer network)
<!-- should use something like "using syntax from Kivela 2014..." or 
Following [@Kivela_2014]:
other info to maybe include: the multilayer network was not node-aligned since
not all individuals were on all layers
the temporal aspect is ordinal coupling and landcover aspect is categorical coupling

@@

#### Network metrics
We used a series of metrics to characterize the multilayer networks focusing on
the relevance of individual layers, similarity of layers and the role of
individuals. We calculated multidegree, the sum of degree centrality across
layers, and neighbourhood, the sum of unique neighbours to each individual
[@Berlingerio_2012; @Kivela_2014]. Comparing the multidegree and number of
unique neighbours, we measured the connective redundancy, the proportion of
neighbours repeated in multiple layers, and the layer relevance, the proportion
of neighbours present compared to the neighbourhood for each individual
[@Berlingerio_2012]. To measure the role of individuals, we calculated the graph
strength for each individual for each layer specific season and landcover
context <!-- citation?-->. Finally, we calculated the similarity of layers
between seasons, across landcover classes (e.g.: winter open compared to summer
open) using the Pearson correlation coefficient [@Br_dka_2018].

### Varying scale in multilayer networks
#### Social scale

@@

#### Spatial scale

@@

#### Temporal scale

@@

## Scale in multilayer networks in practice: case study results and discussion

### Each metric's interpretation
* Multidegree: sum of degree across layers. High multidegree = connected to many
individuals across layers. Low multidegree = not connected to many individuals
across layers.

@@

### Figure captions and brief interpretations

(ref:lcres) Varying scale in landcover resolution. For each landcover resolution (10m, 100m, 250m, 500m and 1000m), multilayer networks were constructed for each landcover class (forest, lichen and open) across two seasons (summer and winter), and four network metrics were calculated. A) Similarity of landcover network layers between seasons (e.g.: open summer compared to open winter) showed different responses to changing landcover scale with an increase in similarity for forest layers at 250m and decrease in similarity for open layers at 250m. Note that lichen layer similarity is not shown because individuals did not interact in lichen habitats greater greater than 10m landcover resolution. B) Mean connective redundancy across individuals showed a decreasing trend with increasing landcover resolution. C) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter lichen layer relevance at higher landcover resolutions. D) Mean graph strength of individuals was weaker during the summer compared to the winter months.

@@

@@

###
Results and interpretation of temporal scale

@@

## Perspectives  to advance multilayer networks 
### Habitat selection and space use

@@

### Individual fitness and phenotypes

@@

### Movement ecology and collective movement

@@

## Conclusions

@@

## Acknowledgements

<!-- add jack -->

We thank all members of the Wildlife Evolutionary Ecology Lab, including K.
Kingdon, S. Boyle,  C. Prokopenko, I. Richmond, J. Hogg, and L. Newediuk for
their comments on previous versions of this manuscript as well as D. C. Schneider
for inspiration and helpful discussions about scale. Funding for this study was
provided by a Vanier Canada Graduate Scholarship to QMRW and a NSERC Discovery
Grant to EVW.

## References
